### 2019-05-31 Output/ModelingUpdates.Rmd
Begin log file for results summaries

---
title: "Demography Model Updates"
author: "Nick Gotelli"
date: ""
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This markdown file gives updates of progress on the demographic modeling. The most recent postings are at the top.

# 31 May 2019

![Caption text can go here.](../Graphics/Figure1.jpeg)

![Caption text here. ](../Graphics/Figure2.jpeg)

![Caption text.](../Graphics/Figure3.jpeg)