### 2019-07-20 Scripts/BromusUpdate.Rmd
Create progress report for Bromus analyses

---
title: "*Bromus* Initial Analysis"
author: "Nick Gotelli"
date: "7/19/2019"
output: 
  html_document: 
    highlight: tango
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data sources and cleaning
These data were provided by Alden Griffith and Michael Loic. After talking with Alden, we decided to use lambda values that had been estimated from changes in the number of individuals in marked plots, rather than analysis of stage transition matrices. For a number of control and snow-manipulated plots, we have the estimated lambda $(N_{t + 1}/N_{t})$ based on annual counts of cheatgrass for 7 consecutive years.

I cleaned the initial data file (`Bromus Lambda and Climate Data.csv`), moving a notes column to initial comment lines, and adding 4 missing values as `NA`. The modified data are in `CleanBromusData.csv`.

```{r}
# Preliminaries -----------------------------------------------------------
library(tidyverse)

# Get input data ----------------------------------------------------------
df <- read.csv(file="../CleanData/CleanBromusData.csv",comment.char="#")
df <- df[,-ncol(df)]                 # remove empty final column
df_clean <- df[complete.cases(df),]  # remove 4 missing lambda

df_preds <- df_clean[,8:18]           # create dataframe of predictor variables
```

## Initial regression models
With the data frame set up, the first analysis was to use each of the predictor variables (snow, degree-days, precipitation, etc.), with $\lambda$ as the response variable. For each variable, I built a linear and a quadratic regression model.

```{r}
# Function Model Fit ------------------------------------------------------
# inputs: response variable, predictor variable
# outputs: adjusted r2 for linear and quadratic regression models

quad_fit <- function(x=df_clean$Snow.Cover, y=df_clean$Lambda){
  M1 <- summary(lm(y~x))
  M2 <- summary(lm(y~poly(x,2)))
  return(c(M1$adj.r.squared,M2$adj.r.squared))
}
# -------------------------------------------------------------------------
# map all predictor variables to quad_fit
summary_df <- map(df_preds,quad_fit,y=df_clean$Lambda)

# put into a summary data frame and add column labels
summary_df <- t(bind_rows(summary_df))
colnames(summary_df) <- c("Linear","Quadratic")
# -------------------------------------------------------------------------
```

## First screening of predictor variables
In this table, each row is a different model with a single predictor variable, and the data frame shows the adjusted r^2^ for each model. The two columns are for a linear model $(y \sim x)$ and a quadratic model $(y \sim x + x^2)$.

```{r}
print(summary_df)
```

A couple things jump out from these univariate regressions. First, the quadratic model almost always fits better than the corresponding linear model (even after imposing a penalty for the additional parameter), so we have good theoretical grounds and empirical support for going with the quadratic models.

I wanted to have predictors related to snow, temperature, and precipitation, as those seem like the best suspects for what is driving distribution. Based on the adjusted r^2^, the best 3 predictors were `Snow.Cover`, `Grow.Season.Precip.Total`, and `Grow.Season.DD.all.days`. There was also a significant effect of `Starting.Density`, which I will discuss near the end.

## Initial multiple regression model

Now taking those 3 best predictors based on snow, precip, and temperature, here is the full model:

```{r}
# best variables are:
# Grow.Season.DD.all.days (4)
# Grow.Season.Precip.Total (5)
# Snow.Cover (6)
best_preds <- df_preds[,4:6]

y <- df_clean$Lambda
x_1 <- best_preds[,1]
x_2 <- best_preds[,2]
x_3 <- best_preds[,3]

model_1 <- summary(lm(y~poly(x_1,2) + poly(x_2,2) + poly(x_3,2)))
print(model_1)

```

## 2-variable model with an interaction term

Although the precipitation variable was initially a strong predictor, its effect is non-significant with both snow cover and degree days in the model, so I dropped it at this point. The next step was to try a 2-variable model but include an interaction term between snow cover and degree days:

```{r}
# Try adding an interaction term -----------------------------------------

model_3 <- summary(lm(y~poly(x_1,2) + poly(x_3,2) + x_1*x_3))
print(model_3)
```

## Final 2-variable prediction model

However, the interaction term does not add anything. So, we are left with just a two-variable quadratic model for predicting $\lambda$:

```{r}
model_2 <- summary(lm(y~poly(x_1,2) + poly(x_3,2)))
print(model_2)

```
In this model, all of the coefficients are significantly different from zero, and the model explains over 50\% of the variance in $\lambda$, which is decent. For now, that is the model we will start with, although it will be easy enough to change if there are other variables you want to consider.

## Functions for fitting and predicting with new values

The next two functions are used to fit the model to the data, and then to predict new values of $\lambda$ based on new inputs of snow cover or degree days.

```{r}
# Function Lambda Model -----------------------------------------------
# inputs: lambda vector, and data frame of predictor variables
# outputs: linear model for forecasting

lambda_model <- function(y=df_clean$Lambda,z=best_preds){
  y <- df_clean$Lambda
  x_1 <- z[,1] # Grow.Season.DD.all.days
  x_2 <- z[,3]  # Snow.Cover
  forecast_model <- lm(y~poly(x_1,2) + poly(x_2,2))
}

# Function Lambda predict -------------------------------------------------
# inputs: linear model, degree days (vector), snow cover (vector)
# outputs: predicted lambda

lambda_predict <- function(m=my_model,degree_days=c(800,600),snow=c(0.83,.74)) {

z <- predict(object=m, newdata=data.frame(x_1=degree_days,x_2=snow))
z[z<0] <- 0   # reset to 0 any predicted negative values for lambda
return(z)
}

```

## Using the functions for forecasting continental distributions

These two functions will make it possible to generate predictions for any values of snow cover and degree days. For example, here the functions are run for two arbitrary vectors of 10 elements each snow cover and degree days. The output is the estimated $\lambda$ for each pair of the variables:

```{r}
lambda_predict(m=lambda_model(),degree_days=sample(600:800,10),snow=runif(10))
```

When we get to the next step, Matt will use this line of code to generate an estimated lambda for any map location for which we have the snow and degree days measurements.

## Sweeping the parameters and constructing a heat map for $\lambda$

Before we build the continental map, it is instructive to sweep the parameters and visualize the response surface of $\lambda$. To set limits on snow cover and degree days, I will consider snow cover between 0 and 100%, and will expand the degree day measurements by 50% beyond the measured limits in Alden and Michael's experiment:

```{r}
# set limits for parameter inputs
snow_low <- 0.0
snow_high <- 1.0
dd_low <- 0.5*min(x_1)
dd_high <- 1.5*max(x_1)

# create parameter data frame with column for output
param_df <- expand.grid(snow_cover=seq(from=snow_low, to=snow_high, length.out=20),
                        degree_days=seq(from=dd_low, to=dd_high,length.out=20))
param_df$lambda <- NA

# apply function to each row
for (i in 1:nrow(param_df)){
  param_df[i,3] <- lambda_predict(m=lambda_model(),
                                  degree_days=param_df[i,2],
                                  snow=param_df[i,1])
}

```

## Problems with the model
Scanning through the completed `param_df` data frame reveals impossibly high values of lambda for most parameter combinations. The problem is coming from the degree days data:

```{r}
qplot(x=best_preds[,1],y=df_clean$Lambda)

```

Because of the cluster of small values at 800 degree days, the quadratic term in the regression model has a positive coefficient, so $\lambda$ increases at both high and low degree days.

I don't see an easy solution, so for now, I am going to drop back to a model just using snow cover, which has a negative coefficient for the quadratic term and therefore decreases lambda at more extreme values.

## Simple snow cover model
So, here are the reworked functions just using snow cover:
```{r}
# Function Lambda Model2-----------------------------------------------
# inputs: lambda vector, and data frame of predictor variables
# outputs: linear model for forecasting

lambda_model2 <- function(y=df_clean$Lambda,z=best_preds){
  y <- df_clean$Lambda
  # x_1 <- z[,1] # Grow.Season.DD.all.days
  x_2 <- z[,3]  # Snow.Cover
  forecast_model <- lm(y~ poly(x_2,2))
}

# Function Lambda predict2 -------------------------------------------------
# inputs: linear model, degree days (vector), snow cover (vector)
# outputs: predicted lambda

lambda_predict2 <- function(m=my_model,snow=c(0.83,.74)) {

z <- predict(object=m, newdata=data.frame(x_2=snow))
z[z<0] <- 0   # reset to 0 any predicted negative values for lambda
return(z)
}

param_df2 <- data.frame(snow_cover=seq(from=snow_low, to=snow_high, length.out=20),
                        lambda=NA)
```

Now with these simplified functions, we run the model once again:

```{r}

# apply function to each row
for (i in 1:nrow(param_df2)){
  param_df2[i,2] <- lambda_predict2(m=lambda_model2(),
                                  snow=param_df2[i,1])
}
plot_1 <- ggplot(param_df2,aes(x=snow_cover,y=lambda)) +
          geom_point() +
          xlab("Snow Cover") +
          ylab("Predicted Lambda")

plot(plot_1)
```

These results look much more reasonable. The maximum predicted $\lambda$ is only 2.5, which is still high, but not unreasonable. And positive population growth ($\lambda > 1$) is predicted for snow cover values from \~ 35% to 99%.

And, for comparison, here is a plot of the original data from which the prediction equation was derived:

```{r}
df <- data.frame(Snow.Cover=df_clean$Snow.Cover, Lambda=df_clean$Lambda)
plot_2 <- ggplot(df,aes(x=Snow.Cover,y=Lambda)) +
          geom_point() +
          xlab("Snow Cover") +
          ylab("Observed Lambda")

plot(plot_2)
```

## Remaining questions and issues

1. Obviously, the biggest problem is what to do about these environmental variables like degree days that do not have well-behaved quadratic functions. I wonder if this is an artifact caused by snow cover being manipulated, but environmental variables not being controlled. However, I thought that Alden had provided me with just data from all the control plots, so I don't know if this is the issue. I was really hoping we could have at least two predictor variables in the regression model, but that may not work for this data set.

2. In order for Matt to use this model for continental projections, we will need snow cover for each pixel in the continental landscape. From talking with Alden, it sounds like those data will be available, although I don't know much about the grain sizes and resolution.

3. In the first set of regression models, there is a strong effect of initial density, suggesting density dependence, but both the linear and quadratic coefficients are positive, so this is actually some form of inverse density dependence. I can't tell if this is an artifact, or if the density effects are what is actually driving the other patterns that we are seeing. But we need to discuss this more.

4. Alden's data set has additional columns indicating snow conditions (Upwind, downwind, plus, minus). I have ignored those designations in this analysis, but they obviously contribute to variation in $\lambda$. If you want, it would be easy enough to go back and subscript the data to fit to fewer points. That will reduce the variance, although I don't think it is going to change the shape of the curves that much.

5. As you can see, the basic model with snow cover looks reasonable, but we need to figure out what, if anything to do with the other environmental variables. Once we get this sorted out and agree on a final regression model, I will clean up the code so that Matt has a single function to work with for generating maps.

6. Michael L. and Alden (and Jen??), we should probably talk again on skype after you have digested these results. I will hold off on any other analyses until we have talked.