### 2019-07-23 Scripts/BromusUpdate.tex
Checking for LaTeX installation

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Bromus Initial Analysis},
            pdfauthor={Nick Gotelli},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{\emph{Bromus} Initial Analysis}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Nick Gotelli}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{7/19/2019}

\begin{document}
\maketitle

\hypertarget{data-sources-and-cleaning}{%
\subsection{Data sources and cleaning}\label{data-sources-and-cleaning}}

These data were provided by Alden Griffith and Michael Loic. After
talking with Alden, we decided to use lambda values that had been
estimated from changes in the number of individuals in marked plots,
rather than analysis of stage transition matrices. For a number of
control and snow-manipulated plots, we have the estimated lambda
\((N_{t + 1}/N_{t})\) based on annual counts of cheatgrass for 7
consecutive years.

I cleaned the initial data file
(\texttt{Bromus\ Lambda\ and\ Climate\ Data.csv}), moving a notes column
to initial comment lines, and adding 4 missing values as \texttt{NA}.
The modified data are in \texttt{CleanBromusData.csv}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Preliminaries -----------------------------------------------------------}
\KeywordTok{library}\NormalTok{(tidyverse)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## -- Attaching packages -------------------- tidyverse 1.2.1 --
\end{verbatim}

\begin{verbatim}
## v ggplot2 3.2.0     v purrr   0.3.2
## v tibble  2.1.3     v dplyr   0.8.3
## v tidyr   0.8.3     v stringr 1.4.0
## v readr   1.3.1     v forcats 0.4.0
\end{verbatim}

\begin{verbatim}
## -- Conflicts ----------------------- tidyverse_conflicts() --
## x dplyr::filter() masks stats::filter()
## x dplyr::lag()    masks stats::lag()
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Get input data ----------------------------------------------------------}
\NormalTok{df <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file=}\StringTok{"../CleanData/CleanBromusData.csv"}\NormalTok{,}\DataTypeTok{comment.char=}\StringTok{"#"}\NormalTok{)}
\NormalTok{df <-}\StringTok{ }\NormalTok{df[,}\OperatorTok{-}\KeywordTok{ncol}\NormalTok{(df)]                 }\CommentTok{# remove empty final column}
\NormalTok{df_clean <-}\StringTok{ }\NormalTok{df[}\KeywordTok{complete.cases}\NormalTok{(df),]  }\CommentTok{# remove 4 missing lambda}

\NormalTok{df_preds <-}\StringTok{ }\NormalTok{df_clean[,}\DecValTok{8}\OperatorTok{:}\DecValTok{18}\NormalTok{]           }\CommentTok{# create dataframe of predictor variables}
\end{Highlighting}
\end{Shaded}

\hypertarget{initial-regression-models}{%
\subsection{Initial regression models}\label{initial-regression-models}}

With the data frame set up, the first analysis was to use each of the
predictor variables (snow, degree-days, precipitation, etc.), with
\(\lambda\) as the response variable. For each variable, I built a
linear and a quadratic regression model.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Function Model Fit ------------------------------------------------------}
\CommentTok{# inputs: response variable, predictor variable}
\CommentTok{# outputs: adjusted r2 for linear and quadratic regression models}

\NormalTok{quad_fit <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(}\DataTypeTok{x=}\NormalTok{df_clean}\OperatorTok{$}\NormalTok{Snow.Cover, }\DataTypeTok{y=}\NormalTok{df_clean}\OperatorTok{$}\NormalTok{Lambda)\{}
\NormalTok{  M1 <-}\StringTok{ }\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(y}\OperatorTok{~}\NormalTok{x))}
\NormalTok{  M2 <-}\StringTok{ }\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(y}\OperatorTok{~}\KeywordTok{poly}\NormalTok{(x,}\DecValTok{2}\NormalTok{)))}
  \KeywordTok{return}\NormalTok{(}\KeywordTok{c}\NormalTok{(M1}\OperatorTok{$}\NormalTok{adj.r.squared,M2}\OperatorTok{$}\NormalTok{adj.r.squared))}
\NormalTok{\}}
\CommentTok{# -------------------------------------------------------------------------}
\CommentTok{# map all predictor variables to quad_fit}
\NormalTok{summary_df <-}\StringTok{ }\KeywordTok{map}\NormalTok{(df_preds,quad_fit,}\DataTypeTok{y=}\NormalTok{df_clean}\OperatorTok{$}\NormalTok{Lambda)}

\CommentTok{# put into a summary data frame and add column labels}
\NormalTok{summary_df <-}\StringTok{ }\KeywordTok{t}\NormalTok{(}\KeywordTok{bind_rows}\NormalTok{(summary_df))}
\KeywordTok{colnames}\NormalTok{(summary_df) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Linear"}\NormalTok{,}\StringTok{"Quadratic"}\NormalTok{)}
\CommentTok{# -------------------------------------------------------------------------}
\end{Highlighting}
\end{Shaded}

\hypertarget{first-screening-of-predictor-variables}{%
\subsection{First screening of predictor
variables}\label{first-screening-of-predictor-variables}}

In this table, each row is a different model with a single predictor
variable, and the data frame shows the adjusted r\textsuperscript{2} for
each model. The two columns are for a linear model \((y \sim x)\) and a
quadratic model \((y \sim x + x^2)\).

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{print}\NormalTok{(summary_df)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                            Linear Quadratic
## Starting.Density          0.37338   0.42397
## Grow.Season.Precip.Snow   0.02884   0.04209
## Grow.Season.Precip.Rain   0.05091   0.39264
## Grow.Season.DD.all.days   0.20921   0.34622
## Grow.Season.Precip.Total  0.09829   0.47520
## Snow.Cover                0.22703   0.45274
## Winter.Precip.Snow        0.05769   0.15705
## Winter.Precip.Rain        0.42714   0.42819
## Winter.DD.no.snowpack     0.00414  -0.00406
## Winter.DD.all.days       -0.00690   0.14229
## Winter.Precip.Total       0.14253   0.28153
\end{verbatim}

A couple things jump out from these univariate regressions. First, the
quadratic model almost always fits better than the corresponding linear
model (even after imposing a penalty for the additional parameter), so
we have good theoretical grounds and empirical support for going with
the quadratic models.

I wanted to have predictors related to snow, temperature, and
precipitation, as those seem like the best suspects for what is driving
distribution. Based on the adjusted r\textsuperscript{2}, the best 3
predictors were \texttt{Snow.Cover}, \texttt{Grow.Season.Precip.Total},
and \texttt{Grow.Season.DD.all.days}. There was also a significant
effect of \texttt{Starting.Density}, which I will discuss near the end.

\hypertarget{initial-multiple-regression-model}{%
\subsection{Initial multiple regression
model}\label{initial-multiple-regression-model}}

Now taking those 3 best predictors based on snow, precip, and
temperature, here is the full model:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# best variables are:}
\CommentTok{# Grow.Season.DD.all.days (4)}
\CommentTok{# Grow.Season.Precip.Total (5)}
\CommentTok{# Snow.Cover (6)}
\NormalTok{best_preds <-}\StringTok{ }\NormalTok{df_preds[,}\DecValTok{4}\OperatorTok{:}\DecValTok{6}\NormalTok{]}

\NormalTok{y <-}\StringTok{ }\NormalTok{df_clean}\OperatorTok{$}\NormalTok{Lambda}
\NormalTok{x_}\DecValTok{1}\NormalTok{ <-}\StringTok{ }\NormalTok{best_preds[,}\DecValTok{1}\NormalTok{]}
\NormalTok{x_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{best_preds[,}\DecValTok{2}\NormalTok{]}
\NormalTok{x_}\DecValTok{3}\NormalTok{ <-}\StringTok{ }\NormalTok{best_preds[,}\DecValTok{3}\NormalTok{]}

\NormalTok{model_}\DecValTok{1}\NormalTok{ <-}\StringTok{ }\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(y}\OperatorTok{~}\KeywordTok{poly}\NormalTok{(x_}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{poly}\NormalTok{(x_}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{poly}\NormalTok{(x_}\DecValTok{3}\NormalTok{,}\DecValTok{2}\NormalTok{)))}
\KeywordTok{print}\NormalTok{(model_}\DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = y ~ poly(x_1, 2) + poly(x_2, 2) + poly(x_3, 2))
## 
## Residuals:
##    Min     1Q Median     3Q    Max 
## -2.114 -0.454 -0.113  0.246  3.516 
## 
## Coefficients:
##               Estimate Std. Error t value Pr(>|t|)
## (Intercept)     1.6351     0.0789   20.72  < 2e-16
## poly(x_1, 2)1   3.1085     1.2259    2.54  0.01254
## poly(x_1, 2)2   4.4275     1.2023    3.68  0.00035
## poly(x_2, 2)1   0.5263     2.0579    0.26  0.79858
## poly(x_2, 2)2   2.0431     1.6022    1.28  0.20478
## poly(x_3, 2)1   5.5628     1.2775    4.35  2.9e-05
## poly(x_3, 2)2  -5.1611     1.8141   -2.84  0.00524
## 
## Residual standard error: 0.879 on 117 degrees of freedom
## Multiple R-squared:  0.612,  Adjusted R-squared:  0.592 
## F-statistic: 30.8 on 6 and 117 DF,  p-value: <2e-16
\end{verbatim}

\hypertarget{variable-model-with-an-interaction-term}{%
\subsection{2-variable model with an interaction
term}\label{variable-model-with-an-interaction-term}}

Although the precipitation variable was initially a strong predictor,
its effect is non-significant with both snow cover and degree days in
the model, so I dropped it at this point. The next step was to try a
2-variable model but include an interaction term between snow cover and
degree days:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Try adding an interaction term -----------------------------------------}

\NormalTok{model_}\DecValTok{3}\NormalTok{ <-}\StringTok{ }\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(y}\OperatorTok{~}\KeywordTok{poly}\NormalTok{(x_}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{poly}\NormalTok{(x_}\DecValTok{3}\NormalTok{,}\DecValTok{2}\NormalTok{) }\OperatorTok{+}\StringTok{ }\NormalTok{x_}\DecValTok{1}\OperatorTok{*}\NormalTok{x_}\DecValTok{3}\NormalTok{))}
\KeywordTok{print}\NormalTok{(model_}\DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = y ~ poly(x_1, 2) + poly(x_3, 2) + x_1 * x_3)
## 
## Residuals:
##    Min     1Q Median     3Q    Max 
## -2.113 -0.392 -0.039  0.297  3.602 
## 
## Coefficients: (2 not defined because of singularities)
##               Estimate Std. Error t value Pr(>|t|)
## (Intercept)    5.19694    3.99197    1.30  0.19550
## poly(x_1, 2)1  8.90080    5.41854    1.64  0.10312
## poly(x_1, 2)2  4.79181    0.97537    4.91  2.9e-06
## poly(x_3, 2)1 20.80155   15.83877    1.31  0.19162
## poly(x_3, 2)2 -4.23724    1.14679   -3.69  0.00033
## x_1                 NA         NA      NA       NA
## x_3                 NA         NA      NA       NA
## x_1:x_3       -0.00691    0.00775   -0.89  0.37398
## 
## Residual standard error: 0.879 on 118 degrees of freedom
## Multiple R-squared:  0.609,  Adjusted R-squared:  0.592 
## F-statistic: 36.8 on 5 and 118 DF,  p-value: <2e-16
\end{verbatim}

\hypertarget{final-2-variable-prediction-model}{%
\subsection{Final 2-variable prediction
model}\label{final-2-variable-prediction-model}}

However, the interaction term does not add anything. So, we are left
with just a two-variable quadratic model for predicting \(\lambda\):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{model_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\KeywordTok{summary}\NormalTok{(}\KeywordTok{lm}\NormalTok{(y}\OperatorTok{~}\KeywordTok{poly}\NormalTok{(x_}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{poly}\NormalTok{(x_}\DecValTok{3}\NormalTok{,}\DecValTok{2}\NormalTok{)))}
\KeywordTok{print}\NormalTok{(model_}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Call:
## lm(formula = y ~ poly(x_1, 2) + poly(x_3, 2))
## 
## Residuals:
##    Min     1Q Median     3Q    Max 
## -2.095 -0.428 -0.078  0.290  3.541 
## 
## Coefficients:
##               Estimate Std. Error t value Pr(>|t|)
## (Intercept)     1.6351     0.0788   20.74  < 2e-16
## poly(x_1, 2)1   4.1393     0.9448    4.38  2.6e-05
## poly(x_1, 2)2   4.8334     0.9734    4.97  2.3e-06
## poly(x_3, 2)1   6.6919     0.9478    7.06  1.2e-10
## poly(x_3, 2)2  -4.7814     0.9704   -4.93  2.7e-06
## 
## Residual standard error: 0.878 on 119 degrees of freedom
## Multiple R-squared:  0.606,  Adjusted R-squared:  0.593 
## F-statistic: 45.8 on 4 and 119 DF,  p-value: <2e-16
\end{verbatim}

In this model, all of the coefficients are significantly different from
zero, and the model explains over 50\% of the variance in \(\lambda\),
which is decent. For now, that is the model we will start with, although
it will be easy enough to change if there are other variables you want
to consider.

\hypertarget{functions-for-fitting-and-predicting-with-new-values}{%
\subsection{Functions for fitting and predicting with new
values}\label{functions-for-fitting-and-predicting-with-new-values}}

The next two functions are used to fit the model to the data, and then
to predict new values of \(\lambda\) based on new inputs of snow cover
or degree days.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Function Lambda Model -----------------------------------------------}
\CommentTok{# inputs: lambda vector, and data frame of predictor variables}
\CommentTok{# outputs: linear model for forecasting}

\NormalTok{lambda_model <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(}\DataTypeTok{y=}\NormalTok{df_clean}\OperatorTok{$}\NormalTok{Lambda,}\DataTypeTok{z=}\NormalTok{best_preds)\{}
\NormalTok{  y <-}\StringTok{ }\NormalTok{df_clean}\OperatorTok{$}\NormalTok{Lambda}
\NormalTok{  x_}\DecValTok{1}\NormalTok{ <-}\StringTok{ }\NormalTok{z[,}\DecValTok{1}\NormalTok{] }\CommentTok{# Grow.Season.DD.all.days}
\NormalTok{  x_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{z[,}\DecValTok{3}\NormalTok{]  }\CommentTok{# Snow.Cover}
\NormalTok{  forecast_model <-}\StringTok{ }\KeywordTok{lm}\NormalTok{(y}\OperatorTok{~}\KeywordTok{poly}\NormalTok{(x_}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{poly}\NormalTok{(x_}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{))}
\NormalTok{\}}

\CommentTok{# Function Lambda predict -------------------------------------------------}
\CommentTok{# inputs: linear model, degree days (vector), snow cover (vector)}
\CommentTok{# outputs: predicted lambda}

\NormalTok{lambda_predict <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(}\DataTypeTok{m=}\NormalTok{my_model,}\DataTypeTok{degree_days=}\KeywordTok{c}\NormalTok{(}\DecValTok{800}\NormalTok{,}\DecValTok{600}\NormalTok{),}\DataTypeTok{snow=}\KeywordTok{c}\NormalTok{(}\FloatTok{0.83}\NormalTok{,.}\DecValTok{74}\NormalTok{)) \{}

\NormalTok{z <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(}\DataTypeTok{object=}\NormalTok{m, }\DataTypeTok{newdata=}\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{x_1=}\NormalTok{degree_days,}\DataTypeTok{x_2=}\NormalTok{snow))}
\NormalTok{z[z}\OperatorTok{<}\DecValTok{0}\NormalTok{] <-}\StringTok{ }\DecValTok{0}   \CommentTok{# reset to 0 any predicted negative values for lambda}
\KeywordTok{return}\NormalTok{(z)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\hypertarget{using-the-functions-for-forecasting-continental-distributions}{%
\subsection{Using the functions for forecasting continental
distributions}\label{using-the-functions-for-forecasting-continental-distributions}}

These two functions will make it possible to generate predictions for
any values of snow cover and degree days. For example, here the
functions are run for two arbitrary vectors of 10 elements each snow
cover and degree days. The output is the estimated \(\lambda\) for each
pair of the variables:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{lambda_predict}\NormalTok{(}\DataTypeTok{m=}\KeywordTok{lambda_model}\NormalTok{(),}\DataTypeTok{degree_days=}\KeywordTok{sample}\NormalTok{(}\DecValTok{600}\OperatorTok{:}\DecValTok{800}\NormalTok{,}\DecValTok{10}\NormalTok{),}\DataTypeTok{snow=}\KeywordTok{runif}\NormalTok{(}\DecValTok{10}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##     1     2     3     4     5     6     7     8     9    10 
## 1.252 2.233 2.048 0.000 2.294 4.602 1.599 0.000 0.000 0.594
\end{verbatim}

When we get to the next step, Matt will use this line of code to
generate an estimated lambda for any map location for which we have the
snow and degree days measurements.

\hypertarget{sweeping-the-parameters-and-constructing-a-heat-map-for-lambda}{%
\subsection{\texorpdfstring{Sweeping the parameters and constructing a
heat map for
\(\lambda\)}{Sweeping the parameters and constructing a heat map for \textbackslash{}lambda}}\label{sweeping-the-parameters-and-constructing-a-heat-map-for-lambda}}

Before we build the continental map, it is instructive to sweep the
parameters and visualize the response surface of \(\lambda\). To set
limits on snow cover and degree days, I will consider snow cover between
0 and 100\%, and will expand the degree day measurements by 50\% beyond
the measured limits in Alden and Michael's experiment:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# set limits for parameter inputs}
\NormalTok{snow_low <-}\StringTok{ }\FloatTok{0.0}
\NormalTok{snow_high <-}\StringTok{ }\FloatTok{1.0}
\NormalTok{dd_low <-}\StringTok{ }\FloatTok{0.5}\OperatorTok{*}\KeywordTok{min}\NormalTok{(x_}\DecValTok{1}\NormalTok{)}
\NormalTok{dd_high <-}\StringTok{ }\FloatTok{1.5}\OperatorTok{*}\KeywordTok{max}\NormalTok{(x_}\DecValTok{1}\NormalTok{)}

\CommentTok{# create parameter data frame with column for output}
\NormalTok{param_df <-}\StringTok{ }\KeywordTok{expand.grid}\NormalTok{(}\DataTypeTok{snow_cover=}\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from=}\NormalTok{snow_low, }\DataTypeTok{to=}\NormalTok{snow_high, }\DataTypeTok{length.out=}\DecValTok{20}\NormalTok{),}
                        \DataTypeTok{degree_days=}\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from=}\NormalTok{dd_low, }\DataTypeTok{to=}\NormalTok{dd_high,}\DataTypeTok{length.out=}\DecValTok{20}\NormalTok{))}
\NormalTok{param_df}\OperatorTok{$}\NormalTok{lambda <-}\StringTok{ }\OtherTok{NA}

\CommentTok{# apply function to each row}
\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\KeywordTok{nrow}\NormalTok{(param_df))\{}
\NormalTok{  param_df[i,}\DecValTok{3}\NormalTok{] <-}\StringTok{ }\KeywordTok{lambda_predict}\NormalTok{(}\DataTypeTok{m=}\KeywordTok{lambda_model}\NormalTok{(),}
                                  \DataTypeTok{degree_days=}\NormalTok{param_df[i,}\DecValTok{2}\NormalTok{],}
                                  \DataTypeTok{snow=}\NormalTok{param_df[i,}\DecValTok{1}\NormalTok{])}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\hypertarget{problems-with-the-model}{%
\subsection{Problems with the model}\label{problems-with-the-model}}

Scanning through the completed \texttt{param\_df} data frame reveals
impossibly high values of lambda for most parameter combinations. The
problem is coming from the degree days data:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{qplot}\NormalTok{(}\DataTypeTok{x=}\NormalTok{best_preds[,}\DecValTok{1}\NormalTok{],}\DataTypeTok{y=}\NormalTok{df_clean}\OperatorTok{$}\NormalTok{Lambda)                                  }
\end{Highlighting}
\end{Shaded}

\includegraphics{BromusUpdate_files/figure-latex/unnamed-chunk-10-1.pdf}

Because of the cluster of small values at 800 degree days, the quadratic
term in the regression model has a positive coefficient, so \(\lambda\)
increases at both high and low degree days.

I don't see an easy solution, so for now, I am going to drop back to a
model just using snow cover, which has a negative coefficient for the
quadratic term and therefore decreases lambda at more extreme values.

\hypertarget{simple-snow-cover-model}{%
\subsection{Simple snow cover model}\label{simple-snow-cover-model}}

So, here are the reworked functions just using snow cover:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Function Lambda Model2-----------------------------------------------}
\CommentTok{# inputs: lambda vector, and data frame of predictor variables}
\CommentTok{# outputs: linear model for forecasting}

\NormalTok{lambda_model2 <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(}\DataTypeTok{y=}\NormalTok{df_clean}\OperatorTok{$}\NormalTok{Lambda,}\DataTypeTok{z=}\NormalTok{best_preds)\{}
\NormalTok{  y <-}\StringTok{ }\NormalTok{df_clean}\OperatorTok{$}\NormalTok{Lambda}
  \CommentTok{# x_1 <- z[,1] # Grow.Season.DD.all.days}
\NormalTok{  x_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\NormalTok{z[,}\DecValTok{3}\NormalTok{]  }\CommentTok{# Snow.Cover}
\NormalTok{  forecast_model <-}\StringTok{ }\KeywordTok{lm}\NormalTok{(y}\OperatorTok{~}\StringTok{ }\KeywordTok{poly}\NormalTok{(x_}\DecValTok{2}\NormalTok{,}\DecValTok{2}\NormalTok{))}
\NormalTok{\}}

\CommentTok{# Function Lambda predict2 -------------------------------------------------}
\CommentTok{# inputs: linear model, degree days (vector), snow cover (vector)}
\CommentTok{# outputs: predicted lambda}

\NormalTok{lambda_predict2 <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(}\DataTypeTok{m=}\NormalTok{my_model,}\DataTypeTok{snow=}\KeywordTok{c}\NormalTok{(}\FloatTok{0.83}\NormalTok{,.}\DecValTok{74}\NormalTok{)) \{}

\NormalTok{z <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(}\DataTypeTok{object=}\NormalTok{m, }\DataTypeTok{newdata=}\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{x_2=}\NormalTok{snow))}
\NormalTok{z[z}\OperatorTok{<}\DecValTok{0}\NormalTok{] <-}\StringTok{ }\DecValTok{0}   \CommentTok{# reset to 0 any predicted negative values for lambda}
\KeywordTok{return}\NormalTok{(z)}
\NormalTok{\}}

\NormalTok{param_df2 <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{snow_cover=}\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from=}\NormalTok{snow_low, }\DataTypeTok{to=}\NormalTok{snow_high, }\DataTypeTok{length.out=}\DecValTok{20}\NormalTok{),}
                        \DataTypeTok{lambda=}\OtherTok{NA}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Now with these simplified functions, we run the model once again:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# apply function to each row}
\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\KeywordTok{nrow}\NormalTok{(param_df2))\{}
\NormalTok{  param_df2[i,}\DecValTok{2}\NormalTok{] <-}\StringTok{ }\KeywordTok{lambda_predict2}\NormalTok{(}\DataTypeTok{m=}\KeywordTok{lambda_model2}\NormalTok{(),}
                                  \DataTypeTok{snow=}\NormalTok{param_df2[i,}\DecValTok{1}\NormalTok{])}
\NormalTok{\}}
\NormalTok{plot_}\DecValTok{1}\NormalTok{ <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(param_df2,}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{snow_cover,}\DataTypeTok{y=}\NormalTok{lambda)) }\OperatorTok{+}
\StringTok{          }\KeywordTok{geom_point}\NormalTok{() }\OperatorTok{+}
\StringTok{          }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Snow Cover"}\NormalTok{) }\OperatorTok{+}
\StringTok{          }\KeywordTok{ylab}\NormalTok{(}\StringTok{"Predicted Lambda"}\NormalTok{)}

\KeywordTok{plot}\NormalTok{(plot_}\DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{BromusUpdate_files/figure-latex/unnamed-chunk-12-1.pdf}

These results look much more reasonable. The maximum predicted
\(\lambda\) is only 2.5, which is still high, but not unreasonable. And
positive population growth (\(\lambda > 1\)) is predicted for snow cover
values from \textasciitilde{} 35\% to 99\%.

And, for comparison, here is a plot of the original data from which the
prediction equation was derived:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{df <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{Snow.Cover=}\NormalTok{df_clean}\OperatorTok{$}\NormalTok{Snow.Cover, }\DataTypeTok{Lambda=}\NormalTok{df_clean}\OperatorTok{$}\NormalTok{Lambda)}
\NormalTok{plot_}\DecValTok{2}\NormalTok{ <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(df,}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{Snow.Cover,}\DataTypeTok{y=}\NormalTok{Lambda)) }\OperatorTok{+}
\StringTok{          }\KeywordTok{geom_point}\NormalTok{() }\OperatorTok{+}
\StringTok{          }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Snow Cover"}\NormalTok{) }\OperatorTok{+}
\StringTok{          }\KeywordTok{ylab}\NormalTok{(}\StringTok{"Observed Lambda"}\NormalTok{)}

\KeywordTok{plot}\NormalTok{(plot_}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{BromusUpdate_files/figure-latex/unnamed-chunk-13-1.pdf}

\hypertarget{remaining-questions-and-issues}{%
\subsection{Remaining questions and
issues}\label{remaining-questions-and-issues}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Obviously, the biggest problem is what to do about these environmental
  variables like degree days that do not have well-behaved quadratic
  functions. I wonder if this is an artifact caused by snow cover being
  manipulated, but environmental variables not being controlled.
  However, I thought that Alden had provided me with just data from all
  the control plots, so I don't know if this is the issue. I was really
  hoping we could have at least two predictor variables in the
  regression model, but that may not work for this data set.
\item
  In order for Matt to use this model for continental projections, we
  will need snow cover for each pixel in the continental landscape. From
  talking with Alden, it sounds like those data will be available,
  although I don't know much about the grain sizes and resolution.
\item
  In the first set of regression models, there is a strong effect of
  initial density, suggesting density dependence, but both the linear
  and quadratic coefficients are positive, so this is actually some form
  of inverse density dependence. I can't tell if this is an artifact, or
  if the density effects are what is actually driving the other patterns
  that we are seeing. But we need to discuss this more.
\item
  Alden's data set has additional columns indicating snow conditions
  (Upwind, downwind, plus, minus). I have ignored those designations in
  this analysis, but they obviously contribute to variation in
  \(\lambda\). If you want, it would be easy enough to go back and
  subscript the data to fit to fewer points. That will reduce the
  variance, although I don't think it is going to change the shape of
  the curves that much.
\item
  As you can see, the basic model with snow cover looks reasonable, but
  we need to figure out what, if anything to do with the other
  environmental variables. Once we get this sorted out and agree on a
  final regression model, I will clean up the code so that Matt has a
  single function to work with for generating maps.
\item
  Michael L. and Alden (and Jen??), we should probably talk again on
  skype after you have digested these results. I will hold off on any
  other analyses until we have talked.
\end{enumerate}

\end{document}